Molecular dynamics simulation of synchronization in driven particles

### 2020-12-24 manuscript/manuscript_ajp_2020.tex
nearly finished with intro

Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been acheived in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.
%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [mor%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%----------------------------------------------------------
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.
%

%external forces
The
dynamics of particles subject to 
an applied external force apply to many physical systems.
For instance the flow of charges in a conductor,
or cells responding to a chemical gradient.
The external force 
increases the diversity of dynamical behaviors,
and can cause particles to flow in
a variety of non-linear complex patterns
that may be 
synchronized.
Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems occurs in mechanical oscillators \cite{chaos}.

%-------------------------------------------------------------

Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.
[elaborate]

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we focus numerical studies 
of the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape. 
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the driven particle.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic

%describe types of physical systems - particle types
In this work,
we perform 
numerical simulations of confined, driven particles
to model a variety of physical phenomena.
%----------------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

>>>>

Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been acheived in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.

%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [mor%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining landscape 
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as charge carries in quantum wires \cite{Tarucha1995} and
slime molds in microchannels \cite{Gholami2015}.
%Biological systems such as
%neuron axons and capillaries can also be studied
%with these models [more detail and references].
%Many such systems execute local oscillations
%about stable points [elaborate].
%----------------------------------------------------------
%The presence of a modulating surface
%can modify these patterns in a variety of ways,
%changing the onset of dynamical flows,
%and the overall flow patterns.
%
%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we focus numerical studies 
of the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

    \bibitem{Tarucha1995}  Tarucha, S.; Honda, T.; Saku, T. Solid State Commun. 1995, 94, 413.

    \bibitem{Gholami2015} A. Gholami, O. Steinbock, V. Zykov, and E. Bodenschatz, Flow-Driven Waves and Phase-Locked Self-Organization in Quasi-One-Dimensional Colonies of Dictyostelium discoideum, Phys. Rev. Lett. {\bf 114}, 018103 (2015).
